Capstone Movielens Report

### 2025-12-30 reports/capstone-movielens.site/05.appendix.Rmd

### Matrix Factorization Functions
\

@@

@@

@@

::: { #func.mf.residual.dataframe.args.file_name .sidebar }
  - **train_set:** ;

@@

@@

@@

::: {.noteblock data-latex=""}
The source code of the [mf.residual.dataframe](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L3) function is defined in the [MF.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L3) script on _GitHub_.
:::
:::

#### [UMGYDE_model.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L14) Function {#func.UMGYDE_model.predict}
\

::: { #func.UMGYDE_model.predict.usage .sidebar }
##### Usage
\

::: { #func.UMGYDE_model.predict.args .sidebar }
##### Arguments
\

::: { #func.UMGYDE_model.predict.args.file_name .sidebar }
  - **test_set:** ;

::: { #func.UMGYDE_model.predict.details .sidebar }
##### Details
\

::: { #func.UMGYDE_model.predict.value .sidebar }
##### Value
\

::: { #func.UMGYDE_model.predict.code .sidebar }
##### Source Code
\

The source code of the [UMGYDE_model.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L14) function is shown below:
```{r eval=FALSE}
UMGYDE_model.predict <- function(test_set) {
  test_set |>
    UMGY_SmoothedDay_effect.predict(rglr.UMGYD_effect)
}
```

::: {.noteblock data-latex=""}
The source code of the [UMGYDE_model.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L14) function is defined in the [MF.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L14) script on _GitHub_.
:::
:::

>>>>

### UMGYDE Model: Matrix Factorization
\

@@

Computes residuals, defined as the difference between the observed and predicted values of the fully trained _UMGYDE Model_ as described in the [User+Movie+Genre+Year+SmoothedDay Effect (UMGYDE) Model] section.

@@

```{r eval=FALSE}
  mf.edx.residual <- mf.residual.dataframe(edx)
```
:::

@@

::: { #func.mf.residual.dataframe.args.train_set .sidebar }
  - **train_set:** Training dataset;

@@

This function calculates the residuals on the *training dataset* passed as the [train_set](#func.mf.residual.dataframe.args.train_set) argument, taking into account all the effects of the previously trained models, and creates a new data frame that includes the results as a calculated column `rsdsl`, along with the `movieId` and `userId` columns selected from the training dataset.
:::

@@

Data frame containing the following three variables:

  - *rsdl*: (Numeric)Calculated column containing results of the function computation;

  - *userId*: (Integer) User Identifier;

  - *movieId* (Integer) Movie Identifier.

@@

::: {.noteblock data-latex=""}
The source code of the [mf.residual.dataframe](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L3) function is also available in the [MF.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/MF.functions.R#L3) script on _GitHub_.
:::
:::

### 2025-12-30 reports/capstone-movielens.site/bak/02.8-1.MF.UMGYDE-model.bak2.Rmd

@@

### MF: Performing Operation

::: {.noteblock data-latex=""}
The complete source code shown in this section is available in the [Perform the Matrix Factorization & Final Test](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2567) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2567) script on _GitHub_.
:::

We assume that the _Matrix Factorization_ method is going to be the last one to provide sufficient results, so we will use the entire `edx` dataset as a _Training Set_ and `final_holdout_test` as a _Test Set_ for our final model, skipping the intermediate _train/test_ operations on the split _train/validation_ datasets.

@@

First, we get residuals from the _UMGYDE Model_ prediction values using the [mf.residual.dataframe](#func.mf.residual.dataframe) function described above:
```{r, eval=FALSE}
  mf.edx.residual <- mf.residual.dataframe(edx)
```
```{r}
  str(mf.edx.residual)
```
#### MF: Transforming Input Data to Be Compatible With the `recosystem` Package
\

Now, let's convert the residuals data and `final_holdout_test` dataset to the input data compatible with the [recosystem](https://cran.r-project.org/web/packages/recosystem/index.html) package to process:
```{r, eval=FALSE}
  set.seed(5430)
  mf.edx.residual.reco <- with(mf.edx.residual, 
                               data_memory(user_index = userId, 
                                           item_index = movieId,
                                           rating = rsdl))

>>>>

::: {.noteblock data-latex=""}
The complete source code shown in this section is available in the [Matrix Factorization (MF)](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2276) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2276) script on _GitHub_.
:::

@@

### MF: Model Building

::: {.noteblock data-latex=""}
The complete source code shown in this section is available in the [Perform the Matrix Factorization & Final Test](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2299) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2299) script on _GitHub_.
:::

We assume that the _Matrix Factorization_ method will be the last one to provide sufficient results, so we will use the entire `edx` dataset as the _Training Set_ and `final_holdout_test` as the _Test Set_ for our final model, skipping the intermediate *training/validation split operations* that were performed for the previous models.

@@

First, we will obtain residuals from the _UMGYDE Model_ prediction values using the following [line of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2314):
```{r main-script@l2314, eval=FALSE}
  mf.edx.residual <- mf.residual.dataframe(edx)
```

::: {.noteblock data-latex=""}
In the code snippets above, we use the [mf.residual.dataframe](#func.mf.residual.dataframe) user-defined described in the [Models Training: Support Functions]/[UMGYDE Model: Utility Functions] section of [Appendix A](#appndx_a).
:::

```{r}
  str(mf.edx.residual)
```

#### MF: Transforming Input Data to Be Compatible With the `recosystem` Package
\

Next, we convert the residuals data and `final_holdout_test` dataset to the input data compatible with the [recosystem](https://cran.r-project.org/web/packages/recosystem/index.html) package using the following [conde snippet](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2319):
```{r main-script@l2319, eval=FALSE}
  set.seed(5430)
  mf.edx.residual.reco <- with(mf.edx.residual, 
                               data_memory(user_index = userId, 
                                           item_index = movieId,
                                           rating = rsdl))
